---
title: "2019.1.24 - Class 2 Bayesian Statistics"
author: "Michael Weisner"
date: "1/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Probability with Discrete Variables

### Sets

+ A set is a collection of elements
+ Elements can be intervals and or isolated elements
+ One often-used set is the set of a real numbers
+ Loosely, real numbers have decimal points
+ Integers are a subset of all real numbers, $ℝ$, denoted $Z$, where the decimal places are .000...
+ Often negative numbers are excluded, eg $ℝ_+$
+ Sets can be categorical
+ In this session we are going to focus on some subset of $Z$

### Random Variables

+ A function is a rule that uniquely maps each element of an input set to some element of an output set
+ A random variable is a FUNCTION from the sample space, $\Omega$, to some subset of $ℝ$ with a probability-based rule

### Sample Space
The sample space, denoted $\Omega$, si the set of all possible outcomes of an observable random variable

LEARN GREEK ALPHABET BETTER!

Six sided dice has $\Omega$ of 1-6

Capital letters, X, indicates a random var and it's lower case indicates a realization of X.


## First Roll in Bowling

+ Each frame in bowling starts with $n = 10$ pins
+ You get 2 rolls per frame to knock down pins
+ First roll's $\Omega$ is 0-10
+ | is read as "given"

Hohn (2009) discussed distributions of the probability of knocking down $X \ge 0$ out of $n \ge X$ pins including $Pr(x | n) = F_x / (-1 + F_{x + 2})$ where $F_x$ is the xth Fibonacci number, ie $F_0 = 1$, $F_1 = 1$, and otherwise $F_x = F_{x-1} + F_{x-2}$

First 13 fibonacci numbers are 1, 1, 2, 3, 45, 8, 13, 21, 34, 55, 89, 144, and 233

## Bowling Function Example
```{r source}
# source("https://tinyurl.com/y9ubz73j")
```
```{r Bowling}
# computes the x-th Fibonacci number without recursion and with vectorization
set.seed(2019124)
F <- function(x) {
  stopifnot(is.numeric(x), all(x == as.integer(x)))
  sqrt_5 <- sqrt(5) # defined once, used twice
  golden_ratio <- (1 + sqrt_5) / 2 # appears with fibonacci sequence
  return(round(golden_ratio ^ (x + 1) / sqrt_5))
}
# probability of knocking down x out of n pins
Pr <- function(x, n = 10) return(ifelse(x > n, 0, F(x) / (-1 + F(n + 2))))

Omega <- 0:10 # 0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
round(c(Pr(Omega), total = sum(Pr(Omega))), digits = 3)

x <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable
```

Above shows the probability of scoring 0 - 10 on the first roll based on the fibonacci sequence

x is a sample of those distributions
```{r}
x
```

What about both rolls?

Define $X_1$ and $X_2$ to respectively be the number of pins knocked down on the first and second rolls of a frame of bowling. What function yields the probability of knocking down $x_2$ pins on your second roll?

$$
Pr(x_2 | X_1 = x_1, n = 10) = F_ ???
$$

```{r}

```


$$PR(x_2 | X_1 = x_1, n = 10)$$ is a **CONDITIONAL** probability

Conditioning is operational...


## From Aristotelian Logic to Bivariate Probability

In R, TRUE = 1 and FALSE = 0 when doing arithmetic operations

```{r}
(TRUE & TRUE) == (TRUE * TRUE)
```

```{r}
(TRUE & FALSE) == (TRUE * FALSE)
```

Can generatlize to probabilities on the $[0,1]$ interval to compute the probability that two or more propostions are simultaneously true

$\cap$ reads as "and" **GENERAL MULTIPLICATION RULE:**
$$Pr(A \cap B) = PR(B) * Pr(A | B) = Pr(A) * Pr(B | A)$$

Loosely, A & B are independent propositions if  A being true or false tells us nothing about hte probability that B is true and vice versa...

Formally, A & B are independent if $Pr(A|B) = PR(A)$ and $Pr(B|A) = Pr(B)$. Thus, $Pr(A \cap B) = Pr(A) * Pr(B)$

**Reasonable Assumptions:**

+ Two rolls in the same frame are not independent
+ Two rolls in diff frames are independent
+ Rolls by two diff people are independent regardless of whether they are in the same frame




