---
title: "GR5065 Assignment 1"
author: "Michael Weisner"
---

# Bowling Revisted

<!-- This is a chunk of R code that gets executed when you "knit" an RMarkdown file -->
<!-- into a HTML or PDF file. You can add other R chunks by clickin on the Insert -->
<!-- menu at the top right. Fill in the answers in the corresponding sections below. -->

```{r}
# computes the x-th Fibonacci number without recursion and with vectorization
F <- function(x) {
  stopifnot(is.numeric(x), all(x == as.integer(x)))
  sqrt_5 <- sqrt(5) # defined once, used twice
  golden_ratio <- (1 + sqrt_5) / 2
  return(round(golden_ratio ^ (x + 1) / sqrt_5))
}

# probability of knocking down x out of n pins
Pr <- function(x, n = 10) return(ifelse(x > n, 0, (F(x) ^ 2) / (F(n) * F(n + 1))))

Omega <- 0:10 # 0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10 is the sample space
```

## Admissibility

#### 1.1 Answer
We know that a Probability Mass Function is admissible because we are dealing with discrete random variables, each with a probability between 0 and 1, and with a cumulative sum of the probablities being 1.

The individiual probabilities of each pin (0:10) are as follows:
```{r}
round(Pr(Omega), digits = 3) 
```

The cumulative probabilities for each pin is as follows:
```{r}
round(cumsum(Pr(Omega)), digits = 3)
```

## Three Cases

```{r}
joint_Pr <- matrix(0, nrow = 11, ncol = 11)
rownames(joint_Pr) <- colnames(joint_Pr) <- as.character(Omega)
for (x1 in Omega) {
  Pr_x1 <- Pr(x1, n = 10)
  for (x2 in 0:(10 - x1)) {
    joint_Pr[x1 + 1, x2 + 1] <- Pr_x1 * Pr(x2, n = 10 - x1)
  }
}
```

#### 1.2.1 Answer

1. The probability of rolling a strike on the first roll of a frame is 0.6181 (as given by the rounded probability of Omega below for hitting 10 pins)

```{r}
round(Pr(Omega), digits = 4)
```

#### 1.2.2 Answer

2. The probability of a spare is a bit more complicated. We need to calculate the probability of each spare combination (0 & 10, 1 & 9, 2 & 8, 3 & 7, 4 & 6, 5 & 5, 6 & 4, 7 & 3, 8 & 2, and 9 & 1).

Using the joint_Pr below, we can see the probability of each spare combination (with the first roll represented by the row and second roll by the column). As each of these combinations are independent, we can get the probability of rolling any spare combination by adding these values. 

```{r}
round(joint_Pr, digits = 5)
```

Let's create a matrix with only the spare possibilities, so dropping the row where we can roll 10 first and dropping the column where we can roll 0 second, since that can only happen in a strike.
```{r}
spares <- joint_Pr[1:10, 2:11] # technically not necessary
round(spares, digits = 5)
```

Now let's add up the diagonals from bottom to top:
```{r}
round(sum(spares[10, 1], spares[9, 2], spares[8, 3], spares[7, 4], spares[6, 5], spares[5, 6], spares[4, 7], spares[3, 8], spares[2, 9], spares[1, 10]), digits = 4) # I couldn't think of an elegant way to do this
```

So the probability of rolling a spare is 0.212

#### 1.2.3 Answer

3.The probability of not getting a strike nor a spare is simply 1 minus the probability of a strike and the probability of a spare, or...
```{r}
1 - 0.6181 - 0.2120
```

0.1699 is the probability of bowling neither a strike nor a spare based on this distribution.

## Conditional Expectation

The expectation of a sum is the sum of the expectations, in this case the expected $x_1 + x_2$.

If we were simply calculating the expected number of pins, then we would calculate the expectation by summing the probabilities of each of the viable frame combination:

```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
sum(joint_Pr * S)
```

But in this case, a spare and strike being non-viable. We can remove those combinations from our sample space like so: 

```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S <- ifelse(S >= 10, 0, S) # makes combinations of 10, which are all strike and spare combos, 0
S
```

And then we can find the sum using the original joint_Pr probabilities like so:
```{r}
sum(joint_Pr * S)
```

#### 1.3 Answer
Which means the expected frame if we roll a non-strike and a non-spare is approximately 1.502 pins.

## Tenth Frame
First we must calculate the probability of an open frame, spare, and strike.

Let's first sum all of the possible non-strike and non spare options of joint_Pr
```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S <- ifelse(S >= 10, 0, S) # makes combinations of 10, which are all strike and spare combos, 0

open_true <- ifelse(S > 0, 1, 0)
open_true[1,1] <- 1 # fix that this is dropped


Pr_open <- sum(joint_Pr * open_true) 
Pr_open
```
We also already calculated the probablity of a spare and a strike
```{r}
Pr_spare <- sum(spares[10, 1], spares[9, 2], spares[8, 3], spares[7, 4], spares[6, 5], spares[5, 6], spares[4, 7], spares[3, 8], spares[2, 9], spares[1, 10])
Pr_strike_open <- Pr(Omega[11]) * (Pr_open + Pr_spare)
Pr_strikex2 <- Pr(Omega[11]) * Pr(Omega[11])

sum(Pr_open, Pr_spare, Pr_strike_open, Pr_strikex2)
```

So..


$\mathbb{E}$[total pins]$= \mathbb{E}$[pins | open frame]$P($open frame$) + \mathbb{E}$[pins | spare frame]$P($spare frame$) + \mathbb{E}$[pins | strike frame]$P($strike frame$)$

So we have to take the expected value of number of pins given the open pins and multiply times the probability of an open frame plus the expected pins of a spare frame times the probability of a spare frame and lastly the expected pins of a strike times the probability of a strike.

Next let's calculate the expectation for each roll.

### Expected Pins for Open Frame
If we bowl an open frame, the expecation, as we calculated before, would be 1.502 pins
```{r}
Ex_open <- sum(joint_Pr * S)
Ex_open
```

### Expected Pins for Spare
If we bowl a spare, then our expected value would be 10
```{r}
Ex_spare <- 10 + sum(Omega * Pr(Omega))
Ex_spare
```


### Expected Pins for One Strike, but No Bonus Strike
And if we get a strike on the first roll of the tenth frame, our expected number of pins would be 10
```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S[11, 1] <- 0 # remove strike possibility of bonus
Ex_strike_open <- 10 + sum(joint_Pr * S) 
Ex_strike_open
```

### Expected pins for One Strike and One Bonus Strike
```{r}
Ex_strike_x2 <- 20 + sum(Omega * Pr(Omega))
Ex_strike_x2
```


$\mathbb{E}$[total pins]$= \mathbb{E}$[pins | open frame]$P($open frame$) + \mathbb{E}$[pins | spare frame]$P($spare frame$) + \mathbb{E}$[pins | strike frame]$P($strike frame$)$

So we have to take the expected value of numb er of pins given the open pins and multiply times the probability of an open frame plus the expected pins of a spare frame times the probability of a spare frame and lastly the expected pins of a strike times the probability of a strike.


So we'll get... 

```{r}
Ex_open * Pr_open + Ex_spare * Pr_spare + Ex_strike_open * Pr_strike_open + Ex_strike_x2 * Pr_strike_x2
```

#### 1.4 Answer
So we can expect roughly 19.041 pins in the final frame.

## Perfect Game

#### 1.5 Answer
This would simply be the probability of striking over 12 frames. Pr(Omega) represents our probabilities on the first roll of a frame, $x_1$ which is the only time a strike is possible. Therefore, we can calculate the probability of a strike by taking the probability of rolling a 10 in our first frame to the power of the number of frames bowled.

```{r}
Pr(Omega[11])^12 # the probability of a strike multiplied for each frame and the 2 additional strikes possible in frame 10
```

Which leaves us with a probability of 0.0031 of rolling a perfect game.

## Scoring

Let's reassert the probabilities of open frames, spares, and strikes.
```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S <- ifelse(S >= 10, 0, S) # makes combinations of 10, which are all strike and spare combos, 0

open_true <- ifelse(S > 0, 1, 0)
open_true[1,1] <- 1 # fix that this is dropped
Pr_open <- sum(joint_Pr * open_true) 
Pr_spare <- sum(spares[10, 1], spares[9, 2], spares[8, 3], spares[7, 4], spares[6, 5], spares[5, 6], spares[4, 7], spares[3, 8], spares[2, 9], spares[1, 10])
Pr_strike_open <- Pr(Omega[11]) * (Pr_open + Pr_spare) # probability of only single strike
Pr_strike_x2 <- Pr(Omega[11]) * Pr(Omega[11]) # probability of double strike

sum(Pr_open, Pr_spare, Pr_strike_open, Pr_strike_x2)
```

Let's also re-establish the expected pins, which should actually double as the score for how I've calculated it.
```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S <- ifelse(S >= 10, 0, S) # makes combinations of 10, which are all strike and spare combos, 0
Ex_open <- sum(joint_Pr * S)

Ex_spare <- 10 + sum(Omega * Pr(Omega))

S <- row(joint_Pr) - 1 + col(joint_Pr) - 1 # for 10th frame
S[11, 1] <- 0 # remove strike possibility of bonus
Ex_strike_open <- 10 + sum(joint_Pr * S) 
Ex_strike_x2 <- 20 + sum(Omega & Pr(Omega))
```


### One Frame Score Expectation
```{r}
Ex_frame <- (Ex_open * Pr_open + Ex_spare * Pr_spare + Ex_strike_open * Pr_strike_open + Ex_strike_x2 * Pr_strike_x2)
Ex_frame
```

So each frame we'd expect a score of approximately 19.041

### Nine Frames Score Expectation
```{r}
Ex_frame * 9
```

Meaning we'd expect a score of 171.39 over 9 frames.

Having done this, I now realize that the expectation of each frame is the same as the tenth score (which was hinted at in the question).

So ultimately for 10 frames of bowling we should expect...

```{r}
Ex_frame * 10
```

#### 1.6 Answer
A score of approximately 190.407.

## Variance

#### 1.7 Answer
The variance of the entire game must be greater than the variance of a single frame, because - while frames are independent from each other - score is potentially dependent on subsequent frames. Also, since the 10th frame has a higher score possibility then the overall variance must be higher than a single frame. 

# Probability in Poker
```{r}
(stacks <- c(Salomon = 27450000, Kaverman = 8025000, Holz = 28675000,
Bonomo = 51150000, Smith = 19700000) / 600000)
```
 
## Going "All In"

#### 2.1 Answer
Since there were only 5 players left, Kaverman was guaranteed a positive outcome, which allowed for some riskier play. The main thing with this situation, was that by going all in on a riskier hand, he still had a chance to win the hand, but going all in may also have caused his opponents to consider him having a stronger hand than he did, so they may have folded and ceded the hand to Kaverman. Since he has the fewest chips, even a small win with his opponents folding could expand his survivability and ultimately position him better for victory of the tournament. Even if his opponents did not fold, he had a statistical justification of going all in - even a weak one - that could have allowed him to greatly increase his chip count. If Kaverman believed his chances of success coupled with the potential gains from winning while going all in outweighed the risk of getting knocked out of the tournament - which still guaranteed him $2 million, then it would be a "logical" choice to go all in.

## Holz's First Decision

### Simulated Method
The probability of any of the 5 cards that are dealt being another 10 would be 2/(52 - 4 cards in their hands) = 2/48 = $\frac {1}{24}$ chance of a 10 being flipped on the first card. Every subsequent dealt card would be the equivalent of 2/48 - n where n is the number of cards dealt. This is assuming we're not counting the 2 cards that were in the other players' hands which could have contained the other pairs.

So, basically, the odds of at least 1 or more 10s being dealt would be $\frac {2}{48} + \frac {2} {47} + \frac {2}{46} + \frac {2}{45} + \frac {2}{44}$ which is equal to

```{r}
(2/48) + (2/47) + (2/46) + (2/45) + (2/44)
```

However, there's an equal probability that one or both of Kaverman's cards are dealt during that time. For simplicity's sake let's assume Kaverman had two aces.

Here are all of the possibilities of the 5 revealed cards:

0 tens drawn, 0 aces : Kaverman Wins
*1 ten, 0 aces: Holz wins*
1 ten, 1 ace: Kaverman wins
1 ten, 2 aces: Kaverman wins
*2 tens, 0 aces: Holz wins*
*2 tens, 1 ace: Holz wins*
2 tens, 2 aces: Kaverman wins

There are three scenarios where Holz can win.

We could simulate this as such:
```{r}
set.seed(232019)
S <- 10^6; win <- 0
deck <- 1:48 # 48 other possible cards left after subtracting the two hands
for (s in 1:S) {
  hand <- sample(deck, 5, replace = FALSE)
if (1 %in% hand & !(2 %in% hand) & !(3 %in% hand) & !(4 %in% hand)) win <- win + 1
if (!(1 %in% hand) & 2 %in% hand & !(3 %in% hand) & !(4 %in% hand)) win <- win + 1
if (1 %in% hand & 2 %in% hand & !(3 %in% hand) & !(4 %in% hand))
if (1 %in% hand & 2 %in% hand & 3 %in% hand & !(4 %in% hand)) win <- win + 1
if (1 %in% hand & 2 %in% hand & !(3 %in% hand) & 4 %in% hand) win <- win + 1
}
(simulated = win / S)
```

Leaving an estimated probability of approximately 0.159 of Holz winning by drawing one or more pairs than Kaverman.

We could also do this with a hypergeometric distribution. The probability of drawing one ten is $\frac {2}{48} = \frac {1}{24}$ and both tens is $\frac {2}{48} * \frac {1}{47} = \frac {1}{1128}$. But we also have to consider that he'll win only if he draws more pairs than Kaverman, who has an equal probability of drawing one or both of his pairs. The probability of Kaverman NOT drawing his pairs is equal to $\frac {46}{48} * \frac {45}{47} = \frac {2025}{2256}$. 

So the outcomes where Holz wins are where he draws one and Kaverman draws none, Holz draws both tens and Kaverman draws none, and Holz draws both tens and Kaverman draws one pair.

This could be thought of as...

 $\frac {2}{48} *  \frac {43}{47} *  \frac {42}{46} * \frac {41}{45} * \frac {40}{44}$
$\frac {2}{48} *  \frac {1}{47} *  \frac {42}{46} * \frac {41}{45} * \frac {40}{44}$
$\frac {2}{48} *  \frac {1}{47} *  \frac {2}{46} * \frac {41}{45} * \frac {40}{44}$
 
### Hypergeometric Method
To figure out these probabilities, we have to use dhyper, I believe separately on both Holz and Kaverman's possibilities.

#### Probability Holz draws 0, 1 or 2 Tens
```{r}
holz_cards <- cbind(0:2, round(dhyper(0:2, 2, 48 - 2, 5) , digits = 6))
holz_cards
```
We can see there's an 0.8 probability that no tens are drawn, a 0.191 probability that 1 is drawn, and approximately a 0.009 probability that both tens are drawn.

This is the same for Kaverman.

Thus, the probability that Holz beats Kaverman would be...

```{r}
holz_cards[2,2] * holz_cards[1,2] + holz_cards[3,2] * holz_cards[1,2] + holz_cards[3,2] * holz_cards[2,2]
```

#### 2.2.1 Answer
**Which gives us the probability that Holz beats Kaverman of 0.161.**


2. If Kaverman has a two cards with different values that are both higher than a ten, what is the probability that Holz wins the hand? Again, for simplicity you can assume that everyone else folds, and ignore the slim possibility that Kaverman or Holz gets a straight, flush, or full house.


### Simulated Method
Let's say this time Kaverman has an Ace and King.

Here are all of the possible solutions where Holz can win:

*0 tens, 0 aces, 0 kings: Holz Wins*
*1 tens, 0 aces, 0 kings: Holz Wins*
*1 tens, 0 ace, 1 kings: Holz Wins*
*1 tens, 1 ace, 0 kings: Holz Wins* 
*1 tens, 1 ace, 1 kings: Holz Wins* # because 3 of a kind beats two pair
*2 tens, 0 aces, 0 kings: Holz Wins*
*2 tens, 0 ace, 1 kings: Holz Wins*
*2 tens, 0 ace, 2 kings: Holz Wins*
*2 tens, 1 ace, 0 kings: Holz Wins*
*2 tens, 1 ace, 1 kings: Holz Wins*
*2 tens, 1 ace, 2 kings: Holz Wins*
*2 tens, 2 ace, 0 kings: Holz Wins*
*2 tens, 2 ace, 1 kings: Holz Wins*

Meaning out of 38 possible pair reveals, Holz has 13 possible winning pairs. Basically, whenever there are more tens than aces and more tens than kings (since 3 kings beats two pairs and 4 kings beats three of a kind and a pair).

Let's simulate it.
```{r}
set.seed(232019)
S <- 10^6; win <- 0
deck <- 1:48 # 48 other possible cards left after subtracting the two hands
for (s in 1:S) {
  hand <- sample(deck, 5, replace = FALSE)
  tens <- 2
  aces <- 1
  kings <- 1
if (1 %in% hand) tens <- tens + 1
if (2 %in% hand) tens <- tens + 1  
if (3 %in% hand) aces <- aces + 1
if (4 %in% hand) aces <- aces + 1
if (5 %in% hand) aces <- aces + 1
if (6 %in% hand) kings <- kings + 1
if (7 %in% hand) kings <- kings + 1
if (8 %in% hand) kings <- kings + 1

if (tens > aces & tens > kings) win <- win + 1
} 
(simulated = win / S)
```

This leaves us with a simulated probability of Holz winning of approximately 0.577.

### Hypergeometric Method
Let's do it again but with the hypergeometric distribution.

While Holz's probability of drawing one or two tens remains the same, Kaverman's distribution is what changes.

Let's look at the probability of Kaverman drawing more aces (since we decided arbitrarily he was looking for aces and kings).
```{r}
kav_cards <- cbind(0:3, round(dhyper(0:3, 3, 48 - 3, 5) , digits = 3))
kav_cards
```

So there's a probability of 0.714 that none of one ace he needs are drawn, a probability of 0.261 that one ace is drawn, a probability of 0.025 that two aces are drawn, and a probability of just under 0.001 that all three are drawn.

He has the same probabilities of 0-3 kings being drawn.


```{r}

holz_cards[1,2] * kav_cards[1,2] * kav_cards[1,2] +
  holz_cards[2,2] * kav_cards[1,2] * kav_cards[1,2] +
  holz_cards[2,2] * kav_cards[1,2] * kav_cards[2,2] +
  holz_cards[2,2] * kav_cards[2,2] * kav_cards[1,2] +
  holz_cards[2,2] * kav_cards[2,2] * kav_cards[2,2] +
  holz_cards[3,2] * kav_cards[1,2] * kav_cards[1,2] +
  holz_cards[3,2] * kav_cards[1,2] * kav_cards[2,2] +
  holz_cards[3,2] * kav_cards[1,2] * kav_cards[3,2] +
  holz_cards[3,2] * kav_cards[2,2] * kav_cards[1,2] +
  holz_cards[3,2] * kav_cards[2,2] * kav_cards[2,2] +
  holz_cards[3,2] * kav_cards[2,2] * kav_cards[3,2] +
  holz_cards[3,2] * kav_cards[3,2] * kav_cards[1,2] +
  holz_cards[3,2] * kav_cards[3,2] * kav_cards[2,2] +
  holz_cards[3,2] * kav_cards[3,2] * kav_cards[3,2]
```

#### 2.2.2 Answer
**Which leaves us with a probability that Holz wins of 0.598.**


3. Based on your answers to the previous two subquestions and additional considerations, explain why the announcer says at 0:54 that “Holz is obviously calling”.

#### 2.2.3 Answer
I think calling is "obvious" in the sense that out of the 1326 possible hands only 24 beat what Holz currently has, indicating he has a pretty strong possibility of winning the hand. He can confidently call the all in because he has a strong chance of winning and can safely suffer a loss if he comes up short

## Holz's Second Decision
*Holz thinks for a while before eventually calling, at which point the video says that Holz has a 50% chance to win the hand, Salomon has a 35% chance, and Kaverman has a 15% chance.*

1. What would their three chances of winning the hand be if Salomon instead had a pair of Aces?

### Hypergeometric Method

We need three different probabilities. We already have Holz's probability of drawing pairs, but need to revise the distribution to be focused on the cards rather than the people and correct the total number being drawn from.

#### Ace Distribution
```{r}
ace_dist <- cbind(0:1, round(dhyper(0:1, 1, 46 - 1, 5) , digits = 6))
ace_dist
```

So there is a probablity of 0.109 that the last ace is drawn.

#### Ten Distribution
```{r}
ten_dist <- cbind(0:2, round(dhyper(0:2, 2, 46 - 2, 5) , digits = 6))
ten_dist
```

#### Fives Distribution
```{r}
five_dist <- cbind(0:3, round(dhyper(0:3, 3, 46 - 3, 5) , digits = 6))
five_dist
```


#### Holz Probability
Holz will win if there are more tens than aces and equal or more tens than fives.
```{r}
pr_holz_win <- ten_dist[2,2] * ace_dist[1,2] * five_dist[1,2] +
  ten_dist[2,2] * ace_dist[1,2] * five_dist[2,2] +
  ten_dist[2,2] * ace_dist[1,2] * five_dist[3,2] +
  ten_dist[3,2] * ace_dist[1,2] * five_dist[1,2] +
  ten_dist[3,2] * ace_dist[1,2] * five_dist[2,2] +
  ten_dist[3,2] * ace_dist[1,2] * five_dist[3,2] +
  ten_dist[3,2] * ace_dist[1,2] * five_dist[4,2] +
  ten_dist[3,2] * ace_dist[2,2] * five_dist[1,2] +
  ten_dist[3,2] * ace_dist[2,2] * five_dist[2,2] +
  ten_dist[3,2] * ace_dist[2,2] * five_dist[3,2] +
  ten_dist[3,2] * ace_dist[2,2] * five_dist[4,2]
pr_holz_win
```

So Holz would have a probability of 0.186 of winning the hand.

#### Salomon Probability
Salomon can win only if the number of aces is equal or greater than tens and fives

```{r}
pr_sal_win <- ace_dist[1,2] * five_dist[1,2] * ten_dist[1,2] +
  ace_dist[1,2] * five_dist[2,2] * ten_dist[1,2] +
  ace_dist[2,2] * five_dist[1,2] * ten_dist[1,2] +
  ace_dist[2,2] * five_dist[1,2] * ten_dist[2,2] +
  ace_dist[2,2] * five_dist[2,2] * ten_dist[1,2] +
  ace_dist[2,2] * five_dist[2,2] * ten_dist[2,2] +
  ace_dist[2,2] * five_dist[3,2] * ten_dist[1,2] +
  ace_dist[2,2] * five_dist[3,2] * ten_dist[2,2]
pr_sal_win
```
So a 0.794 probability of Salomon winning.

#### Kaverman Probability
We could simply calculate this as 1 - the probability of Holz winning - the probability of Salomon winning.
```{r}
pr_kav_win <- 1 - pr_holz_win - pr_sal_win
pr_kav_win
```

So a 0.045 probability of winning. If we wanted to use the distribution we could confirm it with the probability of cutting more fives than both tens and aces.

```{r}
pr_kav_win2 <- five_dist[3,2] * ace_dist[1,2] * ten_dist[1,2] +
  five_dist[4,2] * ace_dist[1,2] * ten_dist[1,2] +
  five_dist[4,2] * ace_dist[1,2] * ten_dist[1,2] +
  five_dist[4,2] * ace_dist[1,2] * ten_dist[2,2] +
  five_dist[4,2] * ace_dist[2,2] * ten_dist[1,2] +
  five_dist[4,2] * ace_dist[2,2] * ten_dist[2,2]
pr_kav_win2
```

I believe the difference above is simply a rounding issue.


#### 2.3.1 Answer

So, the probabilities of winning are as follows:
Holz = 0.186
Salomon = 0.794
Kaverman = 0.020

2. From Holz’s perspective, what is the probability that Salomon has a pair of Aces?

#### 2.3.2 Answer

Technically, since Salomon already has 1 known ace, and Holz knows the two cards that were dealt to himself, there is a 3/49 chance of the reamining card being another ace (49 being all cards that Holz doesn't know). However, since Salomon also went all in Holz would know that Salomon had one of the aforementioned 138 safe hands, of which 15 out of the 138 hands included the Ace of hearts. That being said, only a pair of aces would have been better than Holz's hand, so there was a 3/15 or 1/5 chance to Holz that Salomon had a better hand.

3. Based on your answer’s to the previous two subquestions and additional considerations, was Holz’s decision to call the second time a good decision or a bad decision. Why?

#### 2.3.3 Answer
Given that there was a 1/5 chance that Salomon had a stronger hand, it was still in Holz's favor to call, but those are still very close odds to bet the majority of one's chips on.

## Independent Chip Model
#### 2.4 Answer
*Use the ICM compute the expected amount of prize money Salomon would have won had he not exposed the Ace of hearts.*
```{r}
stacks <- c(Salomon = 45000000, Holz = 20650000, Bonomo = 51150000, Smith = 19700000)
(prob <- stacks / sum(stacks))
```

###**NEED TO FIX **

## Perspectives on Probability

#### 2.5 Answer
Poker involves classical probability because each hand has equal odds of being dealt, frequentist statistics in that you can use behavior like someone going all in to narrow down the possible hand the player has based on those hands being deemed acceptable to safely go in on. Lastly, you could argue it is Bayesian in that you could observe prior hands of an opponent to see how safely or dangerously they play. Perhaps they go in on unsafe hands, or only on very safe hands. These past behaviors can shift your prediction of the player's current behavior.


