---
title: "Scratch Work"
author: "Michael Weisner"
date: "2/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tenth Frame

Hint: Use iterated expectations to obtain a marginal expectation like we did in lecture, where the three marginal probabilities are for a strike, a spare, and an open frame.
```{r}
S <- row(joint_Pr) - 1 + col(joint_Pr) - 1
S
P <- ifelse(S[, 0:10] > 10, 0, S[, 0:10])
P
P <- cbind(P, c(10:20))
P
round(Pr(P), digits = 4)
```




```{r}
S <- 10^6; yes <- 0
for (s in 1:S) {
  x1 <- sample(Omega, size = 1, prob = Pr(Omega))
  x2 <- sample(0:(10 - x1), size = 1, prob = Pr(0:(10 - x1), n = 10 - x1))
  x3 <- sample(0:)
if (x1 == 9 & x2 == 1) yes <- yes + 1
}
c(simulated = yes / S, truth = joint_Pr["9", "1"])
```




```{r}
tenth_Pr <- matrix(0, nrow = 3, ncol = 3)
colnames(tenth_Pr) <- c("Open", "Spare", "Strike")
rownames(tenth_Pr) <- c("X_1", "X_2", "X_3")

for (x1 in Omega) {
  Pr_x1 <- Pr(x1, n = 10)
  for (x2 in 0:(10 - x1)) {
    joint_Pr[x1 + 1, x2 + 1] <- Pr_x1 * Pr(x2, n = 10 - x1)
  }
}
```




If $x_1$ < 10, $x_2$ can be 0:(10-$x_1$) and $x_3$ does not exist and must therefore be 0.
If $x_1$ == 10, $x_2$ can be 0:9 and $x_3$ can be 0:(10-$x_2$)
If $x_1$ == 10 and $x_2$ == 10 then $x_3$ can be 0:10

Total number of possible pins is 30; $x_1 = 10 + x_2 = 10 + x_3 = 10$
Minimum number of possible pins is 0; $x_1 = 0 + x_2 = 0 + x_3 = 0$
If $x_1 = 10$,

0-10 pins is not dependent on anything, 11-20 pins is dependent on x_1 = 10, 20-30 pins is dependent on x_1 = 10 and x_2 = 10.

```{r}
P <- row(joint_Pr) - 1 + col(joint_Pr) - 1
P <- ifelse(P > 10, 0, P)
P
```

```{r}
sum(joint_Pr * P)
```



The expectation of a conditional expectation is a marginal expectation.

x_1 --> marginal probability of a strike is 0.618
x_2 --> given x_1 is a strike: marginal probability
x_2 --> given x_1 is not a strike: marginal probability of a spare is 0.212
x_3 --> given x_1 is not a strike --> 0
x_3 --> given x_1 is a strike and x_2 is a strike: marginal probabilities of omega
x_3 --> given x_1 is not a strike and x_2 is a spare, marginal probabilities of omega 

```{r}
round(rbind(Pr_X1 = Pr(Omega), margin1 = rowSums(joint_Pr), margin2 = colSums(joint_Pr)), 3)

Pr_X1 <- rowSums(joint_Pr) # marginal probabilities of x_1
Pr_X2 <- colSums(joint_Pr) # marginal probabilies of x_2
EX2 <- sum(Omega * Pr_X2)  # definition of marginal expectation
EX2

strike_1 <- Pr(Omega[11])
strike_2 <- Pr(Omega[11]) * Pr(Omega[11])
strike_3 <- Pr(Omega[11]) * Pr(Omega[11]) * Pr(Omega[11])

spare_1 <- Pr()
spare_2 <-
spare_3 <-
  
open_1 <-
open_2 <-
open_3 <-
```  





Simulate scores
```{r}
S <- 1000
for (s in 1:S) {
  x1 <- sample(Omega, size = 1, prob = Pr(Omega))
  x2 <- sample(0:(10 - x1), size = 1, prob = Pr(0:(10 - x1), n = 10 - x1))
  x3 <- sample(Omega, size = 1, prob = Pr(Omega))
  x4 <- sample(0:(10 - x3), size = 1, prob = Pr(0:(10 - x3), n = 10 - x3))
  x5 <- sample(Omega, size = 1, prob = Pr(Omega))
  x6 <- sample(0:(10 - x5), size = 1, prob = Pr(0:(10 - x5), n = 10 - x5))
  x7 <- sample(Omega, size = 1, prob = Pr(Omega))
  x8 <- sample(0:(10 - x7), size = 1, prob = Pr(0:(10 - x7), n = 10 - x7))
  x9 <- sample(Omega, size = 1, prob = Pr(Omega))
  x10 <- sample(0:(10 - x9), size = 1, prob = Pr(0:(10 - x9), n = 10 - x9))
  x11 <- sample(Omega, size = 1, prob = Pr(Omega))
  x12 <- sample(0:(10 - x11), size = 1, prob = Pr(0:(10 - x11), n = 10 - x11))
  x13 <- sample(Omega, size = 1, prob = Pr(Omega))
  x14 <- sample(0:(10 - x13), size = 1, prob = Pr(0:(10 - x13), n = 10 - x13))
  x15 <- sample(Omega, size = 1, prob = Pr(Omega))
  x16 <- sample(0:(10 - x15), size = 1, prob = Pr(0:(10 - x15), n = 10 - x15))
  x17 <- sample(Omega, size = 1, prob = Pr(Omega))
  x18 <- sample(0:(10 - x17), size = 1, prob = Pr(0:(10 - x17), n = 10 - x17))
  x19 <- sample(Omega, size = 1, prob = Pr(Omega))
  x20 <- sample(0:(10 - x19), size = 1, prob = Pr(0:(10 - x19), n = 10 - x19))
  x21 <- sample(Omega, size = 1, prob = Pr(Omega))
  x22 <- sample(0:(10 - x21), size = 1, prob = Pr(0:(10 - x21), n = 10 - x21))
  x23 <- sample(Omega, size = 1, prob = Pr(Omega))
  x24 <- sample(0:(10 - x23), size = 1, prob = Pr(0:(10 - x23), n = 10 - x23))
  x25 <- sample(Omega, size = 1, prob = Pr(Omega))
  x26 <- sample(0:(10 - x25), size = 1, prob = Pr(0:(10 - x25), n = 10 - x25))
  x27 <- sample(Omega, size = 1, prob = Pr(Omega))
  x28 <- sample(0:(10 - x27), size = 1, prob = Pr(0:(10 - x27), n = 10 - x27))
  x29 <- sample(Omega, size = 1, prob = Pr(Omega))
  x30 <- sample(0:(10 - x29), size = 1, prob = Pr(0:(10 - x29), n = 10 - x29))
  x31 <- sample(Omega, size = 1, prob = Pr(Omega))
  x32 <- sample(0:(10 - x31), size = 1, prob = Pr(0:(10 - x31), n = 10 - x31))
  x33 <- sample(Omega, size = 1, prob = Pr(Omega))
  x34 <- sample(0:(10 - x33), size = 1, prob = Pr(0:(10 - x33), n = 10 - x33))
  x35 <- sample(Omega, size = 1, prob = Pr(Omega))
  x36 <- sample(0:(10 - x35), size = 1, prob = Pr(0:(10 - x35), n = 10 - x35))
  x37 <- sample(Omega, size = 1, prob = Pr(Omega))
  x38 <- sample(0:(10 - x37), size = 1, prob = Pr(0:(10 - x37), n = 10 - x37))
  x39 <- sample(Omega, size = 1, prob = Pr(Omega))
  ifelse(x39 < 10, x40 <- sample(0:(10 - x39), size = 1, prob = Pr(0:(10 - x39), n = 10 - x39)), x40 <- sample(Omega, size = 1, prob = Pr(Omega)))
  ifelse((x39 + x40) < 10, x41 <- 0, x41 <- sample(Omega, size = 1, prob = Pr(Omega)))
}
```

```{r}
score <- 0
# Frame 1
  if(x1 + x2 < 10) score <- score + x1 + x2
  if(x1 == 10) score <- score + 10 + x2 + x3
  if(x1 == 10 & x3 == 10 & x5 == 10) score <- score + 30
  if(x1 != 10 & x1 + x2 == 10) score <- score + 10 + x3
score
x1
x2
score

# Frame 2
  if(x3 + x4 < 10) score <- score + x3 + x4
  if(x3 == 10) score <- score + 10 + x5 + x6
  if(x3 == 10 & x5 == 10 & x7 == 10) score <- score + 30
  if(x3 != 10 & x3 + x4 == 10) score <- score + 10 + x5
score
x3
x4
score

# Frame 3
  if(x5 + x6 < 10) score <- score + x5 + x6
  if(x5 == 10) score <- score + 10 + x7 + x8
  if(x5 == 10 & x7 == 10 & x9 == 10) score <- score + 30
  if(x5 != 10 & x5 + x6 == 10) score <- score + 10 + x7
score
x5
x6
score

# Frame 4
  if(x7 + x8 < 10) score <- score + x7 + x8
  if(x7 == 10) score <- score + 10 + x9 + x10
  if(x7 == 10 & x9 == 10 & x11 == 10) score <- score + 30
  if(x7 != 10 & x7 + x8 == 10) score <- score + 10 + x9
score
x7
x8
score

# Frame 5
  if(x9 + x10 < 10) score <- score + x9 + x10
  if(x9 == 10) score <- score + 10 + x11 + x12
  if(x9 == 10 & x11 == 10 & x13 == 10) score <- score + 30
  if(x9 != 10 & x9 + x10 == 10) score <- score + 10 + x11
score
x9
x10
score

# Frame 6
  if(x11 + x12 < 10) score <- score + x11 + x12
  if(x11 == 10) score <- score + 10 + x13 + x14
  if(x11 == 10 & x13 == 10 & x15 == 10) score <- score + 30
  if(x11 != 10 & x11 + x12 == 10) score <- score + 10 + x13
score
x11
x12
score

# Frame 7
  if(x13 + x14 < 10) score <- score + x12 + x14
  if(x13 == 10) score <- score + 10 + x15 + x16
  if(x13 == 10 & x14 == 10 & x15 == 10) score <- score + 30
  if(x13 != 10 & x13 + x14 == 10) score <- score + 10 + x15
score
x13
x14
score

# Frame 8
  if(x15 + x16 < 10) score <- score + x15 + x16
  if(x15 == 10) score <- score + 10 + x17 + x18
  if(x15 == 10 & x16 == 10 & x17 == 10) score <- score + 30
  if(x15 != 10 & x15 + x16 == 10) score <- score + 10 + x17
score
x15
x16
score

# Frame 9
  if(x17 + x18 < 10) score <- score + x17 + x18
  if(x17 == 10 & x19) score <- score + 10 + x19 + x20
  if(x17 == 10 & x18 == 10 & x19 == 10) score <- score + 30
  if(x17 != 10 & x17 + x18 == 10) score <- score + 10 + x19
score
x17
x18
score

# Frame 10
  if(x19 + x20 < 10) score <- score + x19 + x20
  if(x19 == 10) score <- score + 10 + x21 
  if (x19 == 10 & x20 == 10 & x21 < 10) score <- score + 30 + x21
  if(x19 == 10 & x20 == 10 & x21 == 10) score <- score + 30 
  if(x19 != 10 & x19 + x20 == 10) score <- score + 10 + x21
score
x19
x20
score




```


```{r}
S <- 10^6; win <- 0
deck <- 1:48 # 48 cards left
for (s in 1:S) {
  hand <- sample(deck, 5, replace = FALSE)
if (1 %in% hand | 2 %in% hand & !(3 %in% hand) & !(4 %in% hand)) win <- win + 1
if (1 %in% hand & 2 %in% hand & 3 %in% hand & !(4 %in% hand)) win <- win + 1
if (1 %in% hand & 2 %in% hand & !(3 %in% hand) & 4 %in% hand) win <- win + 1
}
(simulated = win / S)
```




### Actual 3 changes of winning
### Simulated Method
```{r}
set.seed(232019)
S <- 10^6; holz <- 0; kav <- 0; sal <- 0
deck <- 1:46 # 46 other possible cards left after subtracting the three hands
for (s in 1:S) {
  hand <- sample(deck, 5, replace = FALSE)
  tens <- 2
  aces <- 1 # because both 
  kings <- 1
  fives <- 1
if (1 %in% hand) tens <- tens + 1
if (2 %in% hand) tens <- tens + 1  
if (3 %in% hand) aces <- aces + 1
if (4 %in% hand) aces <- aces + 1
if (5 %in% hand) kings <- kings + 1
if (6 %in% hand) kings <- kings + 1
if (7 %in% hand) kings <- kings + 1
if (8 %in% hand) fives <- fives + 1
if (9 %in% hand) fives <- fives + 1
if (10 %in% hand) fives <- fives + 1
  
if (tens > aces & tens > kings & tens >= fives) holz <- holz + 1
if (aces >= tens & aces >= kings & aces >= fives) sal <- sal + 1
if (kings > aces & kings >= tens & kings >= fives) sal <- sal + 1
if (fives > aces & fives > kings & fives > tens) kav <- kav + 1
} 
(simulated = c(holz / S, kav / S, sal / S))
```
So Holz has a probability of winning of approximately 0.607 , Kaverman has a probability of winning of 0.240, and Salomon has a probability of winning of 0.369.






